---
title: "Week 2A SQL and R Code Base submission"
author: "Long Lin"
format: html
editor: visual
---

## Week 2A SQL and R: Overview

For the week 2A SQL and R assignment, I gathered movie ratings data from friends (based on a scale of 1-5) on a set of recently popular movies. I then input this data into PostgreSQL. In order to input the data in SQL, I used three tables and a relational model to connect the ratings table with the users and movies. Once the data was in SQL, I used RStudio to process the data with tidyverse.

## Creating the SQL database

First I created a SQL database in PostgreSQL using the following chunk.

```         
CREATE DATABASE "Movie-Ratings"
WITH
OWNER = postgres
ENCODING = 'UTF8'
LC_COLLATE = 'English_United States.1252'
LC_CTYPE = 'English_United States.1252'
LOCALE_PROVIDER = 'libc'
TABLESPACE = pg_default
CONNECTION LIMIT = -1
IS_TEMPLATE = False;

-- Database: Movie-Ratings

-- DROP DATABASE IF EXISTS "Movie-Ratings";
```

Then I created the tables that I will use to store the rating data, I used 3 tables. One table for the users, one table for the movies, and one table for the ratings. I was able to do that with the following code:

```         
CREATE TABLE users (
    user_id SERIAL PRIMARY KEY,
    name VARCHAR(100) NOT NULL
);

CREATE TABLE movies (
    movie_id SERIAL PRIMARY KEY,
    title VARCHAR(100) NOT NULL
);

CREATE TABLE ratings (
    rating_id SERIAL PRIMARY KEY,
    user_id INTEGER NOT NULL REFERENCES users(user_id),
      movie_id INTEGER NOT NULL REFERENCES movies(movie_id),
      rating INTEGER
);
```

## Data

Next I surveyed my friends about some recent movies and what their ratings for the movies were. Here were the results:

Movies:

Oppenheimer, Wicked, Top Gun: Maverick, Zootopia 2, The Housemaid, Captain America: Brave New World

David:

Oppenheimer: 5

Wicked: 3

Top Gun: Maverick: 4

Zootopia 2: 4

Aaron:

Oppenheimer: 4

Top Gun: Maverick: 5

Zootopia 2: 4

The Housemaid: 4

Josh:

Oppenheimer: 4

Wicked: 4

Top Gun: Maverick: 3

Zootopia 2: 3

Cameron:

Wicked: 3

Zootopia 2: 5

Captain America: Brave New World: 4

June:

Top Gun: Maverick: 3

The Housemaid: 5

Captain America: Brave New World: 3

## Inserting Data

Next I inserted the different movies and users into my SQL table using the following query.

```         
INSERT INTO movies (title)
VALUES ('Oppenheimer'), ('Wicked'), ('Top Gun: Maverick'), ('Zootopia 2'), ('The Housemaid'), ('Captain America: Brave New World')
```

```         
INSERT INTO users (name)
VALUES ('David'), ('Aaron'), ('Josh'), ('Cameron'), ('June')
```

Then I added in all the ratings using the following query.

```         
INSERT INTO ratings (user_id, movie_id, rating)
VALUES (1, 1, 5), (1, 2, 3), (1, 3, 4), (1, 4, 4), (1, 5, NULL), (1, 6, NULL),
       (2, 1, 4), (2, 2, NULL), (2, 3, 5), (2, 4, 4), (2, 5, 4), (2, 6, NULL),
       (3, 1, 4), (3, 2, 4), (3, 3, 3), (3, 4, 3), (3, 5, NULL), (3, 6, NULL),  
       (4, 1, NULL), (4, 2, 3), (4, 3, NULL), (4, 4, 5), (4, 5, NULL), (4, 6, 4),
       (5, 1, NULL), (5, 2, NULL), (5, 3, 3), (5, 4, NULL), (5, 5, 5), (5, 6, 3)
```

## R Studio importing

In order to use the data from the SQL Table in R Studio, I imported the data using RPostgres and the tibble library. Then I used dplyr to process the data.

```{r}
require("RPostgreSQL")
library(tibble)
library(dplyr)

con <- dbConnect(dbDriver("PostgreSQL"),
                 dbname = "Movie-Ratings",
                 host = "localhost",
                 port = 5432,
                 user = "***************",
                 password = "***************")


ratings_query <- "SELECT * FROM ratings;"
ratings_data_frame <- dbGetQuery(con, ratings_query)
ratings_data <- as_tibble(ratings_data_frame)

movies_query <- "SELECT * FROM movies;"
movies_data_frame <- dbGetQuery(con, movies_query)
movies_data <- as_tibble(movies_data_frame)

users_query <- "SELECT * FROM users;"
users_data_frame <- dbGetQuery(con, users_query)
users_data <- as_tibble(users_data_frame)

dbDisconnect(con)

```

## Average Ratings

I then used dplyr to find the average ratings given to each movie, not counting the null values for movies that some users had not seen.

```{r}

ratings_data %>%
  group_by(movie_id) %>%
  summarize(
    average_rating = mean(rating, na.rm = TRUE),
    count = n()
  )

```

Next I used dplyr to figure out the number of ratings that each movie got, taking a total count of the ratings based on movie_id and not counting the null values.

```{r}
ratings_data %>%
  group_by(movie_id) %>%
  summarise(
    non_na_count = sum(!is.na(rating))
  )
```

Finally, I used dplyr to figure out the number of ratings that each user gave.

```{r}
ratings_data %>%
  group_by(user_id) %>%
  summarise(
    non_na_count = sum(!is.na(rating))
  )

```
